---
title: "Sep 24, 2019 Mixed Logit not scratch"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mixed Logit not scratch
name: Jikhan Jeong
ref: https://cran.r-project.org/web/packages/mlogit/vignettes/c5.mxl.html (mainly)
ref: https://eml.berkeley.edu/books/choice2.html
ref: Daly, Andrew, Stephane Hess, and Kenneth Train. 2012. “Assuring Finite Moments for Willingness to Pay in Random Coefficient Models.”    
     Transportation 39 (1): 19–31. https://doi.org/10.1007/s11116-011-9331-3.
ref: León, Gianmarco, and Edward Miguel. 2017. “Risky Transportation Choices and the Value of a Statistical Life.” American Economic Journal: 
     Applied Economics 9 (1): 202–28. https://doi.org/10.1257/app.20160140.


```{r}
setwd( "C:/Users/jikhan.jeong/Documents/R/intro econometrics")
getwd()
```

```{r}
library('mlogit')
data("Train", package = "mlogit")

```

# Multinomial Logit
- need to see the data structure

```{r}
Tr <- mlogit.data(Train, shape = "wide", choice = "choice",
varying = 4:11, sep = "_", id.var = "id",
opposite = c("price", "comfort", "time", "change"))
```

```{r}
Tr$price <- Tr$price / 100 * 2.20371
Tr$time <- Tr$time / 60
Train.ml <- mlogit(choice ~ price + time + change + comfort | - 1, Tr) # -1 for without constant
summary(Train.ml)
```

```{r}
coef(Train.ml)[- 1] # without the first coefficient
```
```{r}
coef(Train.ml)[1] # price coefficient
```

We can get the monetary values as follows:

# interpretation :

* We obtain the value of 26 euros for an hour of traveling 
* 5 euros for a change 
* 14 euros to travel in a more comfortable class

```{r}
coef(Train.ml)[- 1] / coef(Train.ml)[1]
```

# Mixed logit : 3 random parameters without correlation

1. time
2. change 
3. comfort

no constnat -1,
using dataset Tr
panel data
random parameter = rpar, n = normal distribution
R = random draw
correlation between random parapmeters
halton method 
optimization method = bhhh

```{r}
Train.mxlu <- mlogit(choice ~ price + time + change + comfort | - 1, Tr,
panel = TRUE, rpar = c(time = "n", change = "n", comfort = "n"), R = 100,  ## n = means normal distribution, rpar = random parameters
correlation = FALSE, halton = NA, method = "bhhh") ## corrleation is not TRUE = no correlation
```

the results contains 3 more variables related to standard deviation of each random coefficient variables

[1] "price" "time" "change" "comfort"
[5] "sd.time" "sd.change" "sd.comfort"

```{r}
summary(Train.mxlu)
```

# Allowing corrleation among the raondom coefficients

 6 parameters which are the elements of the Choleski decomposition of the covariance matrix of the three random parameters.

 [1] "price" "time"
 [3] "change" "comfort"
 [5] "chol.time:time" "chol.time:change"
 [7] "chol.change:change" "chol.time:comfort"
 [9] "chol.change:comfort" "chol.comfort:comfort
 
```{r}
Train.mxlc2 <- update(Train.mxlu, correlation = TRUE) ## corrleation is true
summary(Train.mxlc2)
```

# Preference Space using rpar with model Train.mxlc2 = random coefficient corrleated model

* Marginal utility of time 
 summary(marg.ut.time) displays the unconditional distribution of the marginal utility of time
```{r}
marginal_utility_of_time <- rpar(Train.mxlc2, "time")
summary(marginal_utility_of_time)
```
# WTP space

* the marginal utility by dividing it by the coefficient of a covariate expressed in monetary value (a price for example), 
  taken as a non random parameter. 

* use the norm argument of rpar

The median value (and the mean value as the distribution is symmetric) of transport time is about 33 euros

```{r}
wtp.time <- rpar(Train.mxlc2, "time", norm = "price") # fixed with price, random parameter with time
summary(wtp.time)
```
*In case of correlated random parameters, as the estimated parameters can’t be directly interpreted*

# vcov.mlogit = covariance matrix of correleated variables (default)

```{r}
vcov(Train.mxlc2, what ="rpar")
```

```{r}
summary(vcov(Train.mxlc2, what ="rpar"))
```

# vcov.mlogit = correlation matrix of correleated variables (default)
- standard devations on the diagonal

```{r}
vcov(Train.mxlc2, what = "rpar", type = "cor")
```

```{r}
summary(vcov(Train.mxlc2, what = "rpar", type = "cor"))
```

# the correlation of the correlated random parameters coefficients

```{r}
cor.mlogit(Train.mxlc2)
```

# the covariance of the correlated random parameters

```{r}
cov.mlogit(Train.mxlc2)
```

# Standard deviaton of the corrleated random parameters

```{r}
stdev(Train.mxlc2)
```

# restriction of correlated random parameters

from 3 random parameters(time, change, comfort) to 2 randome paramters (time, change)

*update* using this commend

```{r}
Train.mxlc2_restricted <- update(Train.mxlc2, correlation = c("time", "comfort"))
```

restricted correlated random parameter covariance matrix
```{r}
vcov(Train.mxlc2_restricted, what = "rpar", type = "cov")
```

```{r}
summary(vcov(Train.mxlc2_restricted, what = "rpar", type = "cov"))
```

restricted correlated random parameter covariance matrix
```{r}
vcov(Train.mxlc2_restricted, what = "rpar", type = "cor")
```

```{r}
summary(vcov(Train.mxlc2_restricted, what = "rpar", type = "cor"))
```

# Hypothesis test for the presence of random coefficients and their correlation 

* 3 nested models requires
* 1. no random coefficient
* 2. a random coefficient without corrleation
* 3. a random coefficient with corrleation

# check no correlated random effect
* multinomial vs corrleated random effect model
* Train.ml = multinomial (no random effect)
* Train.mxlc2 = corrleated random effect model

```{r}
library(car) # for hypothesis test
```
LR test
Wald Test
Likelihood test
Score test
```{r}
statpval <- function(x){
    if (inherits(x, "anova")) 
        result <- as.matrix(x)[2, c("Chisq", "Pr(>Chisq)")]
    if (inherits(x, "htest")) result <- c(x$statistic, x$p.value)
    names(result) <- c("stat", "p-value")
    round(result, 3)
}
lr.mxc <- lrtest(Train.mxlc2, Train.ml)
wd.mxc <- waldtest(Train.mxlc2)
lh.mxc <- linearHypothesis(Train.mxlc2, c("chol.time:time = 0",
                                          "chol.time:change = 0", 
                                          "chol.time:comfort = 0", 
                                          "chol.change:change = 0",
                                          "chol.change:comfort = 0", 
                                          "chol.comfort:comfort = 0"))
sc.mxc <- scoretest(Train.ml, 
                    rpar = c(time = "n", change = "n",comfort = "n"), 
                    R = 100, 
                    correlation = TRUE, 
                    halton = NA, 
                    panel = TRUE)

sapply(list(wald = wd.mxc, lh = lh.mxc, score = sc.mxc, lr = lr.mxc),statpval)
```
# Long format data



```{r}
data("ModeCanada", package = "mlogit")
head(ModeCanada)
```


```{r}
class(data("ModeCanada", package = "mlogit"))
```





